---
title: "Statistical Thinking Week 1 Homework"
format: html 
---
```{r}
#| echo: false
suppressMessages(library(tidyverse))
```
# Chapter 2 Questions

**2E1.** Which of the expressions below correspond to the statement: the probability of rain on Monday?

Answer: Pr(rain, Monday)/Pr(Monday)

**2E2.** Which of the following statements correspons to the expression: Pr(Monday|rain)?

Answer: The probability of rain, given that it is Monday

**2E3.** Which of the expressions below correspond to the statement: *the probability that it is Monday,
given that it is raining?*

Answer: Pr(Monday|rain)

# Week 1 Homework

## Question 1
**Suppose the globe tossing data (Lecture 2, Chapter 2) had turned out to be 3 water and 11 land. Construct the posterior distribution.**

```{r}
prosterior_dist = function(p, W, L) {
    return((p^W) * ((1-p)^L))
}

data = tibble(proportion = seq(from = 0, to = 1, length.out = 11))

obs_W = 3
obs_L = 11

data = data %>% 
    mutate(
        ways = prosterior_dist(proportion, obs_W, obs_L)
    )

all_ways = sum(data$ways)

data = data %>% 
    mutate(
        post = round((ways / all_ways), 3)
    )
data
```

```{r}
ggplot(data) +
    geom_line(mapping = aes(x = proportion, y = post))
```

## Question 2
**Using the posterior distribution from 1, compute the posterior predictive distribution for the next 5 tosses of the same globe. I recommend you use the sampling method.**
```{r}
num_tosses = 5

p_samples <- rbeta(1e4, obs_W + 1, obs_L + 1)
W_sim <- rbinom(1e4,size=num_tosses ,p=p_samples)

plot(table(W_sim))
```